---
title: "Integrative Analysis of Two Omics Data Sets"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(plotly)
library(DT)
```

## Introduction
This function is designed to perform integrative analysis of two omics data sets (e.g., m<sup>6</sup>A and RNA-Seq). Taking a quantification matrix as input, deepEA firstly normalize raw quantification based on user-specific normalization method (currently, three normalization methods including **cumulative distribution, Z-score normalization and min-max normalization** are available), then using kmeans clustering method to cluster genes into four groups and visualize genes by an interactive scatter plot.


## A scatter plot showing the differences between two omics data sets
The following scatter plot shows each gene's quantification level in two omics data sets, the colors represents the clusters identified by kmeans algorithm.

```{r echo = FALSE, warning = FALSE, message = FALSE}
cmrMat <- read.table(file = cmrDir,
                     sep = '\t', header = T, quote = "", stringsAsFactors = F)
cmrMat <- na.omit(cmrMat)
cmrMat <- cmrMat[, 1:3]
colnames(cmrMat) <- c("geneID", "cmrLevel", "omics1")
cmrGene <- cmrMat$geneID
method <- "ecdf"
# transform to relative abundance
if(method == "ecdf"){
  cmrMat$cmrLevel.transform <- ecdf(cmrMat$cmrLevel)(cmrMat$cmrLevel)
  cmrMat$omics1.transform <- ecdf(cmrMat$omics1)(cmrMat$omics1)
}else if(method == "z-score"){
  cmrMat$cmrLevel.transform <- (cmrMat$cmrLevel - mean(cmrMat$cmrLevel, na.rm = T))/sd(cmrMat$cmrLevel, na.rm = T)
  cmrMat$omics1.transform <- (cmrMat$omics1 - mean(cmrMat$omics1, na.rm = T))/sd(cmrMat$omics1, na.rm = T)
}else{
  cmrMat$cmrLevel.transform <- (cmrMat$cmrLevel - min(cmrMat$cmrLevel))/(max(cmrMat$cmrLevel) - min(cmrMat$cmrLevel))
  cmrMat$omics1.transform <- (cmrMat$omics1 - min(cmrMat$omics1))/(max(cmrMat$omics1) - min(cmrMat$omics1))
}

kmeans.res <- kmeans(cmrMat[,4:5], as.numeric(k))
cmrMat$cluster <- factor(kmeans.res$cluster, levels = c(1,2,3,4))
if(log == "true"){
	cmrMat$cmrLevel <- log2(cmrMat$cmrLevel + 0.01)
	cmrMat$omics1 <- log2(cmrMat$omics1 + 0.01)
}

library(plotly)
fig <- plot_ly(data = cmrMat, 
               x = ~cmrLevel,
               y = ~omics1,
               color = ~cluster) %>% 
  layout(showlegend = FALSE)
fig

```

## Detailed tabular information
```{r echo = FALSE, warning = FALSE, message = FALSE}
table_options <- function() {
  list(dom = 'Bfrtip',
    pageLength = 8,
    buttons = list(c('copy', 'csv', 'excel', 'pdf')),
    searchHighlight = TRUE,
    colReorder = TRUE,
    scrollX = TRUE,
    fixedColumns = TRUE,
    extensions = 'Responsive',
    deferRender = TRUE,
    scroller = TRUE,
    lengthChange = FALSE
    )
}
colnames(cmrMat) <- c("geneID", "RNA modification level", "Omics 1", "Transformed RNA modification level", "Transformed omics 1", "Cluster")
datatable(
  cmrMat,
  rownames = FALSE,
  editable = TRUE,
  elementId = "linktable",
  class = 'cell-border',
  escape = FALSE,
  options = table_options(),
  extensions = c('Buttons', 'Select')) %>%
  formatStyle(columns = colnames(cmrMat), fontSize = '80%')
```
