---
title: Integrative Analysis of Three Omics Data Sets
output:
  html_document:
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)
library(plotly)
library(DT)
library(dplyr)
library(rmarkdown)
```


```{r echo = FALSE, warning = FALSE, message = FALSE}
source("/home/galaxy/tools/3-ADVANCED-ANALYSIS/Multi-omics_Integrative_Analysis/functions.R")
cmrMat <- read.table(cmrDir, sep = "\t", header = T, quote = "", stringsAsFactors = F)
cmrMat <- na.omit(cmrMat)
colnames(cmrMat) <- c("geneID", "cmrLevel", "omics1", "omics2")
cmrGene <- cmrMat$geneID
# transform to relative abundance
if(method == "ecdf"){
  cmrMat$cmrLevel <- ecdf(cmrMat$cmrLevel)(cmrMat$cmrLevel)
  cmrMat$omics1 <- ecdf(cmrMat$omics1)(cmrMat$omics1)
  cmrMat$omics2 <- ecdf(cmrMat$omics2)(cmrMat$omics2)
}else if(method == "z-score"){
  cmrMat$cmrLevel <- (cmrMat$cmrLevel - mean(cmrMat$cmrLevel, na.rm = T))/sd(cmrMat$cmrLevel, na.rm = T)
  cmrMat$omics1 <- (cmrMat$omics1 - mean(cmrMat$omics1, na.rm = T))/sd(cmrMat$omics1, na.rm = T)
  cmrMat$omics2 <- (cmrMat$omics2 - mean(cmrMat$omics2, na.rm = T))/sd(cmrMat$omics2, na.rm = T)
}else{
  cmrMat$cmrLevel <- (cmrMat$cmrLevel - min(cmrMat$cmrLevel))/(max(cmrMat$cmrLevel) - min(cmrMat$cmrLevel))
  cmrMat$omics1 <- (cmrMat$omics1 - min(cmrMat$omics1))/(max(cmrMat$omics1) - min(cmrMat$omics1))
  cmrMat$omics2 <- (cmrMat$omics2 - min(cmrMat$omics2))/(max(cmrMat$omics2) - min(cmrMat$omics2))
}
```

### Introduction
This function is designed to perform multi-omics integrative analysis. According RNA modifications-related genes' relative abundace in multi-omics, deepEA grouped genes into seven categories. For example, if users would like to integrate m6A with gene expression and translation, the following categories will be illustrated in a ternary plot:

- Balanced
- m<sup>6</sup>A dominant
- m<sup>6</sup>A suppressed
- Expression dominant
- Expression suppressed
- Translation dominant
- Translation suppressed

### Ternary plot showing the differences of each gene in three omics data sets
```{r echo = FALSE, warning = FALSE, message = FALSE}
tt <- as.matrix(cmrMat[,2:4])
res <- apply(tt, 1, defineMod)
cmrMat$mod <- res
cmrMat$mod <- factor(cmrMat$mod, levels = c("balanced", "m6A dominant", "m6A suppressed", "Expression dominant", "Expression suppressed", "Translation dominant", "Translation suppressed"))
p <- plot_ly(
  cmrMat, 
  a = ~cmrLevel, 
  b = ~omics1, 
  c = ~omics2, 
  color = ~mod, 
  type = "scatterternary",
  text = ~geneID
) %>% 
layout(
  ternary = list(
    aaxis = axis("CMR"), 
    baxis = axis("Omics1"), 
    caxis = axis("Omics2")
  )
)
p
```

### Detailed tabular information
```{r echo = FALSE, warning = FALSE, message = FALSE}
table_options <- function() {
  list(dom = 'Bfrtip',
    pageLength = 8,
    buttons = list(c('copy', 'csv', 'excel', 'pdf')),
    searchHighlight = TRUE,
    colReorder = TRUE,
    scrollX = TRUE,
    fixedColumns = TRUE,
    extensions = 'Responsive',
    deferRender = TRUE,
    scroller = TRUE,
    lengthChange = FALSE
    )
}
colnames(cmrMat) <- c("geneID", "RNA modification level", "Omics 1", "Omics 2", "Category")
datatable(
  cmrMat,
  rownames = FALSE,
  editable = TRUE,
  elementId = "linktable",
  class = 'cell-border',
  escape = FALSE,
  options = table_options(),
  extensions = c('Buttons', 'Select')) %>% 
  formatStyle(columns = colnames(cmrMat), fontSize = '80%')
```

### Category differences between duplicated gene pairs
```{r echo = fALSE, warning = FALSE, message = FALSE}
df <- read.table(file = dup_dir,
                 sep = '\t', header = F, quote = '', stringsAsFactors = F)
df$category_m1 <- cmrMat$Category[match(df$V1, cmrMat$geneID)]
df$category_m2 <- cmrMat$Category[match(df$V2, cmrMat$geneID)]

df <- na.omit(df)
library(dplyr)
stat <- df %>% group_by(category_m1, category_m2) %>% summarise(Count = n())
library(ggalluvial)
p <- ggplot(stat,
       aes(axis1 = category_m1, axis2 = category_m2,
           y = Count)) +
  ggalluvial::geom_alluvium(aes(fill = category_m1, color = category_m1),
                            width = 1/12, alpha = 0.8, knot.pos = 0.4) +
  geom_stratum(width = 1/6, color = "grey") +
  geom_label(stat = "stratum", infer.label = TRUE) +
  scale_x_continuous(breaks = 1:2, labels = c("maize1", "maize2")) +
  theme_minimal() +
  scale_color_manual(values = c("#50A683", "#E06B53", "#697C9F", "#BC6597",  "#81BD54", "#E8BD42", "#C8A478")) +
  scale_fill_manual(values = c("#50A683", "#E06B53", "#697C9F", "#BC6597",  "#81BD54", "#E8BD42", "#C8A478")) +
  theme(
    axis.text.x = element_text(size = 12, face = "bold")
  )
p
```

### Category information in duplicated gene pairs
```{r echo = fALSE, warning = FALSE, message = FALSE}
table_options <- function() {
  list(dom = 'Bfrtip',
    pageLength = 8,
    buttons = list(c('copy', 'csv', 'excel', 'pdf')),
    searchHighlight = TRUE,
    colReorder = TRUE,
    scrollX = TRUE,
    fixedColumns = TRUE,
    extensions = 'Responsive',
    deferRender = TRUE,
    scroller = TRUE,
    lengthChange = FALSE
    )
}
colnames(df) <- c("gene 1", "gene 2", "category 1", "category 2")
datatable(
  df,
  rownames = FALSE,
  editable = TRUE,
  elementId = "linktable",
  class = 'cell-border',
  escape = FALSE,
  options = table_options(),
  extensions = c('Buttons', 'Select')) %>%
  formatStyle(columns = colnames(df), fontSize = '80%')
```


